---
title: "Multilevel Meta-Analysis (Pearson R) Tutorial"
author: Hu, M., Koh, P.S., Soh, X.C., Hartanto, A., Majeed, N.M.
output: pdf_document
---

# Introduction

The following script is designed to perform a multilevel meta-analysis using the `metafor` and `lmerTest` package in R.

The analysis is based on data from Lua et al. (2023) and focuses on the relationship between overall relationship between the need for cognition and well-being. The original paper can be found here: https://doi.org/10.1007/s11031-023-10047-w.

The script includes steps for data preparation, effect size calculation, overall effect size computation, forest plot generation, tests for publication bias, and moderation analysis.

The script is structured to be run in RStudio, and it includes comments to guide users through each step of the process.

# Setting Up

This section sets up the working environment, installs necessary packages, loads the required libraries, and reads in the data.

If the `metafor` and `lmerTest` package is not already installed, use the `install.packages()` function to install it.

## Explanation of the Code

-   The `setwd()` function sets the working directory to the location of the script, ensuring that all file paths are relative to the script's location.

-   The `library()` function loads the `metafor` package.

-   The `options()` function is used to adjust the display settings, specifically to disable scientific notation and set the number of digits displayed.

-   The `read.csv()` function reads in the data from a CSV file named "NFCWB.csv", which contains the data drawn from Lua et al.(2024).

```{r}
### Set Up --------------

# Set working directory to that of script's current location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# R version 4.5.0

# Load packages
library(metafor)  # version 4.8-0
library(lmerTest) # version 3.1-3

# Display settings (to disable scientific notation)
options(scipen = 9999, digits = 4)

# Read in data drawn from Lua et al. (2023)
# Original paper: https://doi.org/10.1007/s11031-023-10047-w
mlmmeta_raw = read.csv("NFCWB.csv")
```
# Prepare Data

This section prepares the data for analysis by computing effect sizes for each study and organizing the data frame.

For more information on the `escalc()` function, refer to ?escalc in R.

## Explanation of the Code

-   The `measure` function specifies the type of effect size to be calculated, in this case, "ZCOR" (Fisher's Z). ZCOR is used to calculate the effect size from the raw correlation coefficients instead of Pearson's r because XXX

-   The ri variable `corr_nfcwb` is the column containing the raw correlation coefficients.

-   The ni variable `sample_size` is the column containing the sample sizes for each study.

-   Afterwards, the `escalc()` function computes the effect sizes (yi) and their corresponding sampling variances (vi) for each study.

-   The publication type is categorized into "Published" and "Unpublished" based on the type of publication (e.g., journal article, conference paper, thesis/dissertation).

-   The `mlmmeta` data frame is then sorted by the type of publication (published vs unpublished) to facilitate clearer visualization in the forest plot.

```{r}
### Prepare Data --------------

# Clean data file (reverse correlation for negative well-being)
mlmmeta_raw$corr_nfcwb = with(mlmmeta_raw, ifelse(wellbeing_category == "Negative well-being", -corr_nfcwb, corr_nfcwb))

# Compute effect sizes for each study
mlmmeta = escalc(
  # Type of effect size measure
  measure = "ZCOR",
  
  # Column for raw correlation coefficients
  ri = corr_nfcwb,
  
  # Column for sample sizes
  ni = sample_size,
  
  # Specify data.frame that the information will be extracted from
  data = mlmmeta_raw
  )

# Categorise publication type into "published" and "unpublished"
# Published: Journal articles
# Unpublished: Conference, Panel Data, Thesis/dissertation, Unpublished data
mlmmeta$publication_type = ifelse(
  mlmmeta$publication_type == "Journal article",
  "Published",
  "Unpublished")

# Order the data frame based on publication type and effect sizes (yi)
mlmmeta = mlmmeta[order(mlmmeta$publication_type, mlmmeta$yi), ]
```
# Computing the Overall Effect Size

This section estimates the overall effect size using the `rma.mv()` function from the `metafor` package.

## Explanation of the Multilevel Meta-Analysis Code

-   The `rma.mv()` function is used to compute the overall effect size, accounting for the nested structure of the data.

-   The `random` function specifies the random effects structure, where `~ 1 |  sample_id/meta_id,` indicates that random effects are nested within the sample and meta ID.

-   The `yi` and `vi` functions specify the effect size estimates and their corresponding sampling variances, respectively.

-   The `data` function specifies the data frame containing the effect size estimates and variances.

-   The `summary()` function is used to display the results of the meta-analysis, including the overall effect size estimate and its confidence interval.

-   The `psych::fisherz2r()` function is used to convert the effect size estimates from Fisher's Z to Pearson's r.
```{r}
### Compute Overall Effect Size ------------------- 

# Effect size estimates
mlmmetaresults = rma.mv(
  # Effect size estimates
  yi = yi,
  # Sampling variances
  V = vi,
  # Include random effects for grouping variable (i.e., sample)
  random = ~ 1 | sample_id/meta_id,
  # Specify where to get the data from
  data = mlmmeta
)

# summary function used to provide detailed results of the meta-analysis
summary(mlmmetaresults)

# Convert from Fisher's Z to Pearson's r
mlmmetaresults$b |> psych::fisherz2r() 
```
# Forest Plot

This section generates a forest plot to visually represent the effect sizes and confidence intervals for each study included in the meta-analysis.

The forest plot is created using the `forest()` function from the `metafor` package.

The plot includes the following features:

-   Arrangement of studies by effect sizes

-   Sample size information for both dyslexia and control groups

-   Custom headers for the plot

-   Custom labels for the studies

Saving the forest plot as a PDF file allows for easy sharing and presentation, and allows adjustment to the plot's dimensions.

## Explanation of the Code

-   The `cairo_pdf()` function starts the graphics device driver to create PDF files, and the `file` function specifies the name of the file. Specifically, the `cairo_pdf` function is used for font compatibility.

-   The `width` and `height` function adjust the dimensions of the PDF file.

-   The `forest()` function is used to create the forest plot, and the `tradmetaresults` object contains the results of the meta-analysis.

-   The `order` function specifies the arrangement of studies, with "obs" indicating that the studies should be arranged by effect sizes. To organise by column, replace "obs" with the specific column in the data frame.

-   The `ylim` function sets the y-axis limits for the plot.

-   The `ilab` function is used to add sample size information for both dyslexia and control groups into the forest plot.

-   The `cbind()` function combines the columns indicating the sample size of the groups (dys and control).

-   The `ilab.xpos` function specifies the horizontal arrangement of the columns.

-   The `slab` function is used to label each effect size with its respective study, using the `paste()` function to combine the "Paper" and "Study" columns. The `paste()` function creates the label, and the `sep` function specifies the separator between the columns.

-   The `xlim` function sets the x-axis limits for the plot.

-   The `alim` function sets the confidence interval limits, and the `steps` function determines the number of intervals in the x-axis.

-   The `efac` function changes the size of effect size polygons.

-   The `header` function is set to FALSE to allow for manual specification of headers.

-   The `xlab` function specifies the confidence interval label for the funnel plot, in this case, "Fisher's Z"

-   The `text()` function is used to manually include text within the plot, such as the "Author(s) Year" header and specific sample size column headers.

-   The `x` and `y` function in the `text()` function adjust the position of the headers, with the `x` function specifying the horizontal arrangement of the columns and the `y` function specifying the vertical arrangement of the columns.

-   The `x` function specifies the horizontal arrangement of the columns and the `y` function specifies the vertical arrangement of the columns.

-   The `font` function adjusts the font size.

-   The `pos` function in the `text()` function specifies the position of the text relative to the specified coordinates.

-   The `font` function adjusts the font size.

-   The `dev.off()` function is used to close the graphics device and finalize the plot as a saved file.
```{r}
### Forest Plot --------------

# Save the forest plot as a PDF file
# Name the pdf file of the forest plot
# cairo_pdf function used for font compatibility
# Adjust the width and height of the pdf file
cairo_pdf(file = "NFCWBforestplot.pdf", width = 14, height = 35)

# Start creating the forest plot itself
# Specify dataset
forest(
  mlmmetaresults, 
  
  # Arrangement of studies
  order = "obs",
  
  # Add y-axis limits
  ylim = c(-3, 111),
  
  # Add sample size information for presence and absence of smartphones group
  # Values indicate the x-axis position of the sample size columns  
  ilab = sample_size,
  ilab.xpos = -3,
  
  # Label studies on the forest plot
  slab = paste(author, year, sep = ", "),
  
  # Add x-axis limits
  xlim = c(-5, 3),
  
  # Add confidence interval limits
  # Adjust intervals based on the number of steps
  alim = c(-1.5, 1.5),
  steps = 7,
  
  # Change size of effect size polygons
  efac = 0.3,
  
  # Show (TRUE) or hide (FALSE) default headers
  # Hide when we want to manually specify our own headers
  header = FALSE,
  
  # Add label for confidence interval, in this case, "Fisher's Z"
  xlab = "Fisher's Z"
)

# For the following lines of code,
# Use text function to manually include text within the plot

# Add "Author(s) Year" header
text(x = -4.6, y = 110, "Author(s) Year", font = 2)

# Add "Sample Size" header
text(x = -3, y = 110, "Sample Size", font = 2)

# Add "r [95% CI]" header
text(x = 2.7, y = 110, "Z [95% CI]", font = 2) 

# Close the forest plot and finalise it as a saved file
dev.off()
```

```{r}
### Tests for Publication Bias --------------------

# Funnel Plot #

# Save the funnel plot as a PDF file
# Name the pdf file of the funnel plot
# Adjust the width and height of the pdf file
pdf(file = "NFCWBfunnelplot.pdf", width = 8, height = 5)
# funnel function to create the funnel plot, specify the data to create the plot
funnel(mlmmetaresults, legend = TRUE, xlab = "Fisher's Z")
# Close the funnel plot and finalise it as a saved file
dev.off()
```

```{r}
### Tests for Publication Bias --------------------

# Rank Correlation Test #
ranktest(mlmmetaresults)
```

```{r}
### Tests for Publication Bias --------------------

# Eggers' Test # 
lmer(
  # g weighted by SE is predicted by intercept and inverse SE
  # with random intercept by sample
  I(yi / vi) ~ 1 + I(1 / vi) + (1 | sample_id),
  data = mlmmeta
) |>
  # Estimate of interest is the intercept
  summary(correlation = TRUE)
```

```{r}
### Moderation Analysis --------------

# Categorical Variable (i.e., publication type)
rma.mv(
  yi = yi,
  V = vi,
  random = ~ 1 | sample_id/meta_id,
  # Specify categorical moderator (i.e., Published articles)
  subset = (publication_type == "Published"),
  data = mlmmeta,
  # To address convergence issues (if it exists)
  control=list(rel.tol=1e-8) 
)

rma.mv(
  yi = yi,
  V = vi,
  random = ~ 1 | sample_id/meta_id,
  # Specify categorical moderator (i.e., Unpublished articles)
  subset = (publication_type == "Unpublished"),
  data = mlmmeta,
  # To address convergence issues (if it exists)
  control=list(rel.tol=1e-8) 
)
```

