---
title: "Traditional Meta-Analysis Tutorial"
output: Traditional Meta-Analysis_notebook
---
The following script is designed to perform a traditional meta-analysis using the `metafor` package in R. 
The analysis is based on data from Majeed et al. (2021) and focuses on the relationship between dyslexia and creativity. The original paper can be found [here](https://doi.org/10.1002/dys.1677).
The script includes steps for data preparation, effect size calculation, overall effect size computation, forest plot generation, tests for publication bias, and moderation analysis.
The script is structured to be run in RStudio, and it includes comments to guide users through each step of the process.

# Setting Up
This section sets up the working environment, installs necessary packages, loads the required libraries, and reads in the data.
```{r}
# Set working directory to that of script's current location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# R version 4.4.3

# Install packages (if not already installed)
install.packages("metafor")

# Load packages
library("metafor") # version 4.8-0

# Display settings (to disable scientific notation)
options(scipen = 9999, digits = 4)

# Read in data drawn from Majeed et al. 2021
tradmeta_raw = read.csv("DYSCRE.csv")
```

# Prepare Data
This section prepares the data for analysis by computing effect sizes and organizing the data frame.
The data frame is sorted by the type of creativity measure and corresponding effect sizes to facilitate clearer visualization in the forest plot.
The data frame does not include pre-computed effect sizes for each study. Therefore, the `escalc()` function from `metafor` package  is used to calculate them. 
For more information on the `escalc()` function, refer to ?escalc in R.
```{r}
tradmeta = escalc(
  # Type of effect size measure
  measure = "SMD",

  # Columns for sample size of each group
  n1i = n_dys,
  n2i = n_control,

  # Columns for means of each group
  m1i = Mean_CRE_dys,
  m2i = Mean_CRE_control,

  # Columns for standard deviation of each group
  sd1i = SD_CRE_dys,
  sd2i = SD_CRE_control,

  # Specify data.frame that the information will be extracted from
  data = tradmeta_raw
)

# Convert Creativity.Measure_type to a factor with specified levels
tradmeta$Creativity.Measure_type = factor(
  tradmeta$Creativity.Measure_type,
  levels = c("Verbal", "Mixed", "Non-verbal")
)
# Order the data frame by Creativity.Measure_type and effect sizes (yi)
tradmeta = tradmeta[order(tradmeta$Creativity.Measure_type, tradmeta$yi), ]
```

# Computing the Overall Effect size
This section estimates the overall effect size using the `rma()` function from the `metafor` package, based on the computed effect sizes (yi) and their corresponding sampling variances (vi).
The `rma()` function is used to perform a random-effects meta-analysis, which accounts for the variability between studies.
For more information on the `rma()` function, refer to ?rma in R.
The method used is REML (Restricted Maximum Likelihood), which is a common approach for estimating the overall effect size in meta-analysis.
The results are summarized to provide detailed information about the meta-analysis.
```{r} 
# Estimate the overall effect size using the rma() function
tradmetaresults = rma(
  # Effect size estimates
  yi = yi,
  # Sampling variance
  vi = vi,
  # Specify method to estimate heterogeneity
  method = "REML",
  # Specify where to get the data from
  data = tradmeta
)

# summary function used to provide detailed results of the meta-analysis
summary(tradmetaresults)
```

# Forest Plot
This section generates a forest plot to visually represent the effect sizes and confidence intervals for each study included in the meta-analysis. 
The forest plot is created using the `forest()` function from the `metafor` package.
The plot includes the following features:
- Arrangement of studies by effect sizes
- Sample size information for both dyslexia and control groups
- Custom headers for the plot
- Custom labels for the studies
Saving the forest plot as a PDF file allows for easy sharing and presentation, and allows adjustment to the plot's dimensions.

## Explanation of the Code  
The `pdf()` function starts the graphics device driver to create PDF files, and the `file` argument specifies the name of the file.
The `width` and `height` arguments adjust the dimensions of the PDF file.
The `forest()` function is used to create the forest plot, and the `tradmetaresults` object contains the results of the meta-analysis.
The `order` argument specifies the arrangement of studies, with "obs" indicating that the studies should be arranged by effect sizes. 
To organise by column, replace "obs" with the specific column in the data frame.
The `ylim` argument sets the y-axis limits for the plot.
The `ilab` argument is used to add sample size information for both dyslexia and control groups into the forest plot.
The `cbind()` function combines the columns indicating the sample size of the groups (dys and control).
The `ilab.xpos` argument specifies the horizontal arrangement of the columns, while the `xlim` argument sets the x-axis limits for the plot.
the `slab` argument is used to label each effect size with its respective study, using the `paste()` function to combine the "Paper" and "Study" columns.
The `paste()` function creates the label, and the `sep` argument specifies the separator between the columns.
The `alim` argument sets the confidence interval limits, and the `steps` argument determines the number of intervals in the x-axis.
The `header` argument is set to FALSE to allow for manual specification of headers.
The `text()` function is used to manually include text within the plot, such as the "Author(s) Year" header and specific sample size column headers.
The `x` and `y` arguments in the `text()` function adjust the position of the headers, while the `font` argument adjusts the font size.
The `dev.off()` function is used to close the graphics device and finalize the plot as a saved file.

```{r}
### Forest Plot --------------

# Save the forest plot as a PDF file
# Name the file of the forest plot
pdf(file = "tradforestplot.pdf", width = 11, height = 9)

# Start creating the forest plot itself
# Specify dataset
forest(
  tradmetaresults,

  # Arrangement of studies
  order = "obs",

  # Add y-axis limits
  ylim = c(-2, 16),

  # Add sample size information for dyslexia (n_dys) and control (n_control) group into forest plot
  ilab = cbind(n_dys, n_control),
  ilab.xpos = c(-4.2, -3.8),

  # Label studies on the forest plot
  slab = paste(Paper, paste("Study", Study), sep = ", "),

  # Add x-axis limits
  xlim = c(-8, 4),

  # Add confidence interval limits
  # Adjust intervals based on the number of steps
  alim = c(-2.5, 2.5),
  steps = 11,

  # Show (TRUE) or hide (FALSE) default headers
  # Hide when we want to manually specify our own headers
  header = FALSE
)

# For the following lines of code,
# use text function to manually include text within the plot

# Add "Author(s) Year" header
text(x = -7.2, y = 14.5, "Author(s) Year", font = 2)

# Add “Sample Size” header
text(x = -4.0, y = 15, "Sample Size", font = 2) # text function includes text within the plot

# Add specific sample size column headers, “Dslx” (Dyslexia Group) and “Ctrl” (Control Group)
# x values represents the x-coordinates of where the "Dslx" and "Ctrl" headers will be placed 
text(c(x = -4.2, x = -3.8), y = 14.5, c("Dslx", "Ctrl"), font = 2)

# Add "g [95% CI]" header
text(x = 3.5, y = 14.5, "g [95% CI]", font = 2) # text function includes text within the plot

# Close the forest plot and finalise it as a saved file
dev.off()
```

# Tests for Publication Bias 
This section performs tests for publication bias, including a funnel plot, rank correlation test and Egger's test.
The funnel plot visually represents the distribution of effect sizes and their standard errors, allowing for the identification of potential publication bias.
Saving the funnel plot as a PDF file allows for easy sharing and presentation, and allows adjustment to the plot's dimensions.
The rank correlation test serves as a complementary method to the funnel plot in assessing publication bias by examining the correlation between effect sizes and their standard errors.
The Egger's test is a statistical test that quantifies the degree of asymmetry in the funnel plot, providing a more formal assessment of publication bias.

## Explanation of the Code
The `pdf()` function starts the graphics device driver to create PDF files, and the `file` argument specifies the name of the file.
The `width` and `height` arguments adjust the dimensions of the PDF file.
The `funnel()` function is used to create the funnel plot, and the `tradmetaresults` object contains the results of the meta-analysis.
The `legend` argument specifies whether to include a legend in the plot.
The `ranktest()` function computes the Kendall tau value, which indicates the strength and direction of the association between the ranks of effect sizes and their standard errors.
The `sei_corrected` variable is calculated to adjust the standard error for the effect sizes, and the `rma()` function is used to perform Egger's test.
The `mods` argument specifies the moderator variable, which is the corrected standard error in this case.
The `weights` argument specifies the weight for each effect size, which is the inverse of the corrected standard error squared.
The `data` argument specifies the dataset to be used for the analysis.
The `summary()` function provides the results of the Egger's test, including the slope estimate and its significance.
The `rma()` function is used to perform Egger's test, which assesses the presence of publication bias by examining the relationship between effect sizes and their standard errors.
The `summary()` function provides the results of the Egger's test, including the slope estimate and its significance.


```{r}
# Funnel Plot

# pdf function starts the graphics device driver to create PDF files
# Name the file of the funnel
# Adjust the width and height of the pdf file
pdf(file = "tradfunnelplot.pdf", width = 8, height = 5)
# funnel argument to create the funnel plot, specify the data to create the plot
funnel(tradmetaresults, legend = TRUE)
# Close the funnel plot and finalise it as a saved file
dev.off()

# Rank Correlation Test

# ranktest argument to compute kendall tau value
ranktest(tradmetaresults)

# Egger's Test

# Calculate standard error (SE)
tradmeta$sei_corrected = with(
  tradmeta,
  sqrt((n_dys + n_control) / (n_dys * n_control))
)
rma(
  # Effect size estimates
  yi = yi,
  # Sampling variance
  vi = vi,
  # Indicate moderator which is SE/sei_corrected
  mods = ~sei_corrected,
  # Indicate weight which is inverse SE^2
  weights = 1 / sei_corrected^2,
  # Specify dataset
  data = tradmeta
) |>
  # Estimate of interest is the slope
  summary()
```

# Moderation Analysis
This section performs moderation analysis to explore the influence of different moderators on the effect sizes. It includes both continuous and categorical moderators.
```{r}
# Continuous variable (i.e., female proportion)
rma(
  yi = yi,
  vi = vi,
  # Specify continuous moderator (i.e., sex)
  mods = ~Proportion.of.female,
  method = "REML",
  data = tradmeta
)

# Categorical variable (i.e., type of creativity measure)
rma(
  yi = yi,
  vi = vi,
  # Specify categorical moderator (i.e., verbal)
  subset = (Creativity.Measure_type == "Verbal"),
  method = "REML",
  data = tradmeta
)
rma(
  yi = yi,
  vi = vi,
  # Specify categorical moderator (i.e., mixed)
  subset = (Creativity.Measure_type == "Mixed"),
  method = "REML",
  data = tradmeta
)
rma(
  yi = yi,
  vi = vi,
  # Specify categorical moderator (i.e., non-verbal)
  subset = (Creativity.Measure_type == "Non-verbal"),
  method = "REML",
  data = tradmeta
)
```

# Forest Plot with Moderators
This section generates a forest plot that includes moderators to visually represent the effect sizes and confidence intervals for each study included in the meta-analysis. 
The forest plot includes the following features:
- Arrangement of studies by effect sizes and types of moderators
- Sample size information for both dyslexia and control groups
- Custom headers for the plot
- Custom labels for the studies 
- Summary effect sizes for each moderator
The plot is saved as a PDF file for easy sharing and presentation.

```{r}
# Name the file of the forest plot
# Adjust the width and height of the pdf file
pdf(file = "tradforestplotwithmoderators.pdf", width = 11, height = 10) # pdf function starts the graphics device driver to create PDF files
forest(
  tradmetaresults,

  # Manually arrange effect sizes by creativity measure type
  # - Mixed: Rows 20 to 18
  # - Verbal: Rows 14 to 10
  # - Non-verbal: Rows 6 to 2
  # The arrangement must consider spacing and must end at row 2
  rows = c(20:18, 14:10, 6:2),

  # Add y-axis limits
  ylim = c(-2, 24),

  # Add sample size information for dyslexia (n_dys) and control (n_control) group into forest plot
  ilab = cbind(n_dys, n_control),
  ilab.xpos = c(-3.8, -3.3),

  # Label studies on the forest plot
  slab = paste(Paper, paste("Study", Study), sep = ", "),

  # Add x-axis limits
  xlim = c(-7, 4),

  # Add confidence interval limits
  # Adjust intervals based on the number of steps
  alim = c(-2.5, 2.5),
  steps = 11,

  # Remove headers (if any), for manual input
  header = FALSE
)

# Add text labels for moderator (type of creativity task)
# Adjust the position of the labels with x (horizontal) and y (vertical) function
# Labels for different creativity task types (Moderator Analysis)
# - "Verbal" at y = 7
# - "Mixed" at y = 15
# - "Non-verbal" at y = 21
# `pos = 4` ensures left alignment of the text
text(
  x = -7,
  pos = 4,
  y = c(7, 15, 21),
  c("Verbal", "Mixed", "Non-verbal"),
  font = 2
)

# Moderation analysis
# subset argument ensures only the relevant rows are used
res.v <- rma(
  yi,
  vi,
  subset = (Creativity.Measure_type == "Verbal"),
  data = tradmeta
)
res.n <- rma(
  yi,
  vi,
  subset = (Creativity.Measure_type == "Non-verbal"),
  data = tradmeta
)
res.m <- rma(
  yi,
  vi,
  subset = (Creativity.Measure_type == "Mixed"),
  data = tradmeta
)

# Add summary effect sizes for each of the moderators
# addpoly argument adds a summary effect size (diamond shape) for each moderator
# row argument places the summary at the corresponding position in the plot
addpoly(res.v, row = 1) # summary effect for "verbal" group
addpoly(res.n, row = 9) # summary effect for "non-verbal" group
addpoly(res.m, row = 17) # summary effect for "mixed" group

# Add"Author(s) Year" header
text(x = -6.3, y = 23, "Author(s) Year", font = 2) # text function includes text within the plot

# Add “Sample Size” header
text(x = -3.6, y = 23.8, "Sample Size", font = 2)

# Add specific sample size column headers, “Dslx” (Dyslexia Group) and “Ctrl” (Control Group)
text(x = c(-3.8, -3.3), y = 23, c("Dslx", "Ctrl"), font = 2)

# Add "g [95% CI]" header
text(x = 3.5, y = 23, "g [95% CI]", font = 2) # text function includes text within the plot

# Close the forest plot and finalise it as a saved file
dev.off()
```