---
title: "Traditional Meta-Analysis Tutorial"
output: Traditional Meta-Analysis_notebook
---

# Setting Up

```{r}
# Set working directory to that of script's current location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# R version 4.4.3

# Install packages (if not already installed)
install.packages("metafor")

# Load packages
library("metafor") # version 4.8-0

# Display settings (to disable scientific notation)
options(scipen = 9999, digits = 4)

# Read in data drawn from Majeed et al. 2021
# Original paper: https://doi.org/10.1002/dys.1677
tradmeta_raw = read.csv("DYSCRE.csv")
```

# Prepare Data

```{r}
# escalc function is used to compute effect sizes
tradmeta = escalc(
  # Type of effect size measure
  # See ?escalc for more information
  measure = "SMD",

  # Columns for sample size of each group
  n1i = n_dys,
  n2i = n_control,

  # Columns for means of each group
  m1i = Mean_CRE_dys,
  m2i = Mean_CRE_control,

  # Columns for standard deviation of each group
  sd1i = SD_CRE_dys,
  sd2i = SD_CRE_control,

  # Specify data.frame that the information will be extracted from
  data = tradmeta_raw
)

# Fix row order of data frame
# For easier plotting later
# Convert Creativity.Measure_type to a factor with specified levels
tradmeta$Creativity.Measure_type = factor(
  tradmeta$Creativity.Measure_type,
  levels = c("Verbal", "Mixed", "Non-verbal")
)
# Order data frame based on type of creativity measure, and based on effect sizes, within each type of creativity measure
tradmeta = tradmeta[order(tradmeta$Creativity.Measure_type,tradmeta$yi), ]
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

```{r}
 
```

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
