---
title: "Multilevel Meta-Analysis Tutorial"
author: Hu, M., Koh, P.S., Soh, X.C., Hartanto, A., Majeed, N.M.
output: pdf_document
---

# Introduction

The following script is designed to perform a multilevel meta-analysis using the `metafor` and `lmerTest` package in R.

The analysis is based on data from Hartanto et al. (2024) and focuses on the relationship between smartphone presence and cognitive functions. The original paper can be found here: <https://doi.org/10.1037/tmb0000123>.

The script includes steps for data preparation, effect size calculation, overall effect size computation, forest plot generation, tests for publication bias, and moderation analysis.

The script is structured to be run in RStudio, and it includes comments to guide users through each step of the process.

# Setting Up

This section sets up the working environment, installs necessary packages, loads the required libraries, and reads in the data.

If the `metafor` and `lmerTest` package is not already installed, use the `install.packages()` function to install it.

## Explanation of the Code

-   The `setwd()` function sets the working directory to the location of the script, ensuring that all file paths are relative to the script's location.

-   The `library()` function loads the `metafor` package.

-   The `options()` function is used to adjust the display settings, specifically to disable scientific notation and set the number of digits displayed.

-   The `read.csv()` function reads in the data from a CSV file named "SPC.csv", which contains the data drawn from Hartanto et al. (2024).

-   A new column named "ID" is created in the data frame to assign unique IDs to each row of data.

```{r}
### Set Up --------------

# R version 4.5.0

# Set working directory to that of script's current location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Load packages
library(metafor)  # version 4.8-0
library(lmerTest) # version 3.1-3

# Display settings (to disable scientific notation)
options(scipen = 9999, digits = 4)

# Read in data drawn from Hartanto et al. 2024
multilevelmeta_raw = read.csv("SPC.csv")

# Create new column with unique IDs
multilevelmeta_raw$ID = 1:nrow(multilevelmeta_raw)
```

# Prepare Data

This section prepares the data for analysis by computing effect sizes for each study and organizing the data frame.

The `multimeta_raw` data frame does not include pre-computed effect sizes for each study. Therefore, the `escalc()` function from `metafor` package is used to calculate them.

For more information on the `escalc()` function, refer to ?escalc in R.

## Explanation of the Code

-   The `measure` function specifies the type of effect size to be calculated, in this case, "SMD" (Standardized Mean Difference).

-   The `n1i` and `n2i` function specify the columns for the sample sizes of each group

-   The `m1i` and `m2i` function specify the columns for the means of each group.

-   The `sd1i` and `sd2i` function specify the columns for the standard deviations of each group.

-   Afterwards, the `escalc()` function computes the effect sizes (yi) and their corresponding sampling variances (vi) for each study.

-   The `multilevelmeta$publication` variable is converted to a factor with specified levels to ensure that the creativity measure types are ordered correctly in the forest plot.

-   The `multilevelmeta` data frame is then sorted by the type of creativity measure and corresponding effect sizes to facilitate clearer visualization in the forest plot.

```{r}
### Prepare Data --------------
# Compute effect sizes for each study
multilevelmeta = escalc(
       # Type of effect size measure
       measure = "SMD",

       # Columns for sample size of each group
       n1i = n_p,
       n2i = n_a,

       # Columns for means of each group
       m1i = cog_M_p,
       m2i = cog_M_a,

       # Columns for standard deviation of each group
       sd1i = cog_SD_p,
       sd2i = cog_SD_a,

       # Specify data.frame that the information will be extracted from
       data = multilevelmeta_raw
)

# Convert Publication to a factor with specified levels
multilevelmeta$publication = factor(
       multilevelmeta$publication,
       levels = c("Journal article", "Thesis/dissertation", "Conference")
)
# Order the data frame based on publication and effect sizes (yi)
multilevelmeta = multilevelmeta[order(multilevelmeta$publication, multilevelmeta$yi), ]
```

# Computing the Overall Effect Size

This section estimates the overall effect size using the `rma.mv()` function from the `metafor` package.

Specifically, we demonstrate the difference in calculating the overall effect size using a traditional meta-analysis approach versus a multilevel meta-analysis approach.

The traditional meta-analysis approach is not recommended for multilevel data, as it does not account for the nested structure of the data (i.e., studies nested within labs).

The multilevel meta-analysis approach is more appropriate for this type of data, as it accounts for the hierarchical structure and allows for the inclusion of random effects.

## Explanation of the Multilevel Meta-Analysis Code

-   The `rma.mv()` function is used to compute the overall effect size, accounting for the nested structure of the data.

-   The `random` function specifies the random effects structure, where `~ 1 | lab_id / ID` indicates that random intercepts are included for both the lab and individual studies.

-   The `yi` and `vi` functions specify the effect size estimates and their corresponding sampling variances, respectively.

-   The `data` function specifies the data frame containing the effect size estimates and variances.

-   The `summary()` function is used to display the results of the meta-analysis, including the overall effect size estimate and its confidence interval.

```{r}
### Compute Overall Effect size --------------

# Using traditional meta code (refer to meta_traditional.R)
# The code below is for demonstration purposes only
# The code below is not recommended for multilevel meta-analysis
usingsimplemetacode = rma(
       yi = yi,
       vi = vi,
       data = multilevelmeta
)
summary(usingsimplemetacode)

# Using multilevel code
# The code below is the recommended approach for multilevel meta-analysis
mlmmetaresults = rma.mv(
       # Effect size estimates
       yi = yi,
       # Sampling variances
       V = vi,
       # Include random effects for grouping variable (i.e., lab)
       random = ~ 1 | lab_id / ID,
       # Specify where to get the data from
       data = multilevelmeta
)

# summary function used to provide detailed results of the meta-analysis
summary(mlmmetaresults)
```

# Forest Plot

This section generates a forest plot to visually represent the effect sizes and confidence intervals for each study included in the meta-analysis.

The forest plot is created using the `forest()` function from the `metafor` package.

The plot includes the following features:

-   Arrangement of studies by effect sizes

-   Sample size information for both presence of smartphones and absence of smartphones groups

-   Custom headers for the plot

-   Custom labels for the studies

Saving the forest plot as a PDF file allows for easy sharing and presentation, and allows adjustment to the plot's dimensions.

## Explanation of the Code

-   The `pdf()` function starts the graphics device driver to create PDF files, and the `file` function specifies the name of the file.

-   The `width` and `height` function adjust the dimensions of the PDF file.

-   The `forest()` function is used to create the forest plot, and the `tradmetaresults` object contains the results of the meta-analysis.

-   The `order` function specifies the arrangement of studies, with "obs" indicating that the studies should be arranged by effect sizes. To organise by column, replace "obs" with the specific column in the data frame.

-   The `ylim` function sets the y-axis limits for the plot.

-   The `ilab` function is used to add sample size information for presence of smartphones and absence of smartphones groups into the forest plot.

-   The `cbind()` function combines the columns indicating the sample size of the groups (dys and control).

-   The `ilab.xpos` function specifies the horizontal arrangement of the columns.

-   The `slab` function is used to label each effect size with its respective study, using the `paste()` function to combine the "Paper" and "Study" columns. The `paste()` function creates the label, and the `sep` function specifies the separator between the columns.

-   The `xlim` function sets the x-axis limits for the plot.

-   The `alim` function sets the confidence interval limits, and the `steps` function determines the number of intervals in the x-axis.

-   The `efac` function changes the size of effect size polygons.

-   The `header` function is set to FALSE to allow for manual specification of headers.

-   The `xlab` function specifies the confidence interval label for the funnel plot, in this case, "Hedge's g"

-   The `text()` function is used to manually include text within the plot, such as the "Author(s) Year" header and specific sample size column headers.

-   The `x` and `y` function in the `text()` function adjust the position of the headers, with the `x` function specifying the horizontal arrangement of the columns and the `y` function specifying the vertical arrangement of the columns.

-   The `x` function specifies the horizontal arrangement of the columns and the `y` function specifies the vertical arrangement of the columns.

-   The `font` function adjusts the font size.

-   The `pos` function in the `text()` function specifies the position of the text relative to the specified coordinates.

-   The `font` function adjusts the font size.

-   The `dev.off()` function is used to close the graphics device and finalize the plot as a saved file.

```{r}
### Forest Plot --------------

# Save the forest plot as a PDF file
# Name the pdf file of the forest plot
pdf(file = "mlmforestplot.pdf", width = 15, height = 40)

# Start creating the forest plot itself
# Specify dataset
forest(
        mlmmetaresults,

       # Arrangement of studies
       order = "obs",

       # Add y-axis limits
       ylim = c(-3, 140),

       # Add sample size information for presence of smartphones and absence of smartphones groups
       # -3 for presence of smartphones (n_p)
       # -2.55 for absence of smartphones (n_a)
       ilab = cbind(n_p, n_a),
       ilab.xpos = c(-3, -2.55),

       # Label studies on the forest plot
       slab = paste(author, year_published, sep = ", "),

       # Add x-axis limits
       xlim = c(-5, 3),

       # Add confidence interval limits
       # Adjust intervals based on the number of steps
       alim = c(-2, 2),
       steps = 9,

       # Change size of effect size polygons
       efac = 0.3,

       # Show (TRUE) or hide (FALSE) default headers
       # Hide when we want to manually specify our own headers
       header = FALSE,

       # Add label for confidence interval, in this case, "Hedge's g"
       xlab = "Hedge's g"
)

# For the following lines of code,
# Use text function to manually include text within the plot

# Add "Author(s) Year" header
text(x = -4.6, y = 139, "Author(s) Year", font = 2)

# Add "Sample Size" header
text(x = -2.8, y = 139.6, "Sample Size", font = 2)

# Add specific sample size column headers, “Presence” and “Absence”
# x = -3 for for presence of smartphones
# x = -2.55 for absence of smartphones
# y = 139 for both
text(c(x = -3, x = -2.55), y = 139, c("Presence", "Absence"), font = 2)

# Add "g [95% CI]" header
text(x = 2.7, y = 139, "g [95% CI]", font = 2) 

# Close the forest plot and finalise it as a saved file
dev.off()
```

# Tests for Publication Bias

This section performs tests for publication bias, including a funnel plot and Egger's test.

For multilevel meta-analysis, we do not recommend conducting a rank correlation test as it is prone to Type 1 error.

Researchers may refer to this article by Fernández-Castilla et al. (2019) on a detailed discussion of the limitations of rank correlation tests in multilevel meta-analysis, as well as an overview on other publication bias tests: <https://doi.org/10.1080/00220973.2019.1582470>:

## Funnel Plot

The funnel plot visually represents the distribution of effect sizes and their standard errors, allowing for the identification of potential publication bias. Saving the funnel plot as a PDF file allows for easy sharing and presentation, and allows adjustment to the plot's dimensions.

### Explanation of the Code

-   The `pdf()` function starts the graphics device driver to create PDF files, and the `file` function specifies the name of the file.

-   The `width` and `height` function adjust the dimensions of the PDF file.

-   The `funnel()` function is used to create the funnel plot, and the `tradmetaresults` object contains the results of the meta-analysis.

-   The `legend` function specifies whether to include a legend in the plot. TRUE indicates that a legend should be included, FALSE indicates that it should not.

-   The `xlab` function specifies the confidence interval label for the funnel plot, in this case, "Hedge's g"

-   The `dev.off()` function is used to close the graphics device and finalize the plot as a saved file.

```{r}
### Tests for Publication Bias --------------

# Funnel Plot #

# Save the funnel plot as a PDF file
# Name the pdf file of the funnel plot
# Adjust the width and height of the pdf file
pdf(file = "mlmfunnelplot.pdf", width = 8, height = 5)
# funnel function to create the funnel plot, specify the data to create the plot
funnel(mlmmetaresults, legend = TRUE, xlab = "Hedge's g")
# Close the funnel plot and finalise it as a saved file
dev.off()
```

## Egger's Test

The Egger's test is a statistical test that quantifies the degree of asymmetry in the funnel plot, providing a more formal assessment of publication bias.

### Explanation of the Code

-   The `multilevelmeta$sei_corrected` variable is created to store the corrected standard error for each effect size, calculated using the formula $\sqrt{\frac{(n_{p} + n_{a})}{(n_{p} \cdot n_{a})}}$. Using Egger' test (unadjusted) on SMDs results in inflated type 1 error as SMD and SE are not independent. Hence, use corrected formula for SE.

-   The `lmer()` function is used to fit a linear mixed-effects model, where the effect size weighted by the corrected standard error is predicted by the intercept and the inverse of the corrected standard error. metafor::rma.mv does not have a weights argument, and metafor::regtest does not support rma.mv objects. For three (or more) level meta-analysis, use `lmerTest::lmer` instead.

-   The `I(yi / sei_corrected)` expression indicates that the effect size (yi) is divided by the corrected standard error (sei_corrected), which is used to weight the effect sizes in the model.

-   The `I(1 / sei_corrected)` expression indicates that the inverse of the corrected standard error is included in the model as a predictor.

-   The `1 | lab_id` expression indicates that random intercepts are included for each lab, accounting for the nested structure of the data.

-   The `data` function specifies the dataset to be used for the analysis.

-   The `summary()` function provides the results of the Egger's test, including the slope estimate and its significance.

```{r}
### Tests for Publication Bias --------------

# Egger' Test #

# Calculate standard error (SE)
multilevelmeta$sei_corrected = with(
       multilevelmeta, 
       sqrt((n_p + n_a) / (n_p * n_a)))
lmer(
       # g weighted by SE is predicted by intercept and inverse SE
       # with random intercept by sample
       I(yi / sei_corrected) ~ 1 + I(1 / sei_corrected) + (1 | lab_id),
       data = multilevelmeta
) |>
       # Estimate of interest is the intercept
       summary(correlation = FALSE)
```

# Moderation Analysis

This section performs moderation analysis to explore the influence of categorical and continuous moderators on the effect sizes.

## Explanation of the Code

-   The `rma.mv()` function is used to compute the multilevel meta-analysis with moderators.

-   The `yi` and `vi` functions specify the effect size estimates and their corresponding sampling variances, respectively.

-   The `random` function specifies the random effects structure, where `~ 1 | lab_id / ID` indicates that random intercepts are included for both the lab and individual studies, accounting for the nested structure of the data.

-   The `mods` function specifies the moderator variable, which can be either continuous or categorical.

-   The `method` function specifies the method used to estimate heterogeneity, in this case, "REML" (Restricted Maximum Likelihood).

-   The `subset` function is used to specify the subset of data for each moderator analysis.

-   The `data` function specifies the data frame containing the effect size estimates and variances.

-   The `summary()` function is used to display the results of the meta-analysis, including the overall effect size estimate and its confidence interval.

Note that if convergence issues arise, the `control` function can be used to address it. Researchers may also refer more to the `metafor` package documentation for more information on how to address convergence issues.

```{r}
### Moderation Analysis --------------

# Categorical variable (i.e., publication)
rma.mv(
       yi = yi,
       V = vi,
       random = ~ 1 | lab_id / ID,
       # Specify categorical moderator (i.e., Journal Article)
       subset = (publication == "Journal article"),
       data = multilevelmeta,
       # To address convergence issues (if it exists)
       control=list(rel.tol=1e-8) 
)
rma.mv(
       yi = yi,
       V = vi,
       random = ~ 1 | lab_id / ID,
       # Specify categorical moderator (i.e., Thesis/dissertation)
       subset = (publication == "Thesis/dissertation"),
       data = multilevelmeta
)
rma.mv(
       yi = yi,
       V = vi,
       random = ~ 1 | lab_id / ID,
       # Specify categorical moderator (i.e., Conference)
       subset = (publication == "Conference"),
       data = multilevelmeta
)

# Continuous variable (i.e., female proportion)
rma.mv(
       yi = yi,
       V = vi,
       random = ~ 1 | lab_id / ID,
       # Specify continuous moderator (i.e., female proportion)
       mods = ~female_proportion,
       method = "REML",
       data = multilevelmeta
) |>
       summary()
```

# Forest Plot of Moderators

This section generates a forest plot that includes moderators to visually represent the effect sizes and confidence intervals for each study included in the meta-analysis.

The plot is saved as a PDF file for easy sharing and presentation.

The forest plot includes the following features:

-   Arrangement of studies by effect sizes and types of moderators

-   Sample size information for both presence of smartphones and absence of smartphones groups

-   Custom headers for the plot

-   Custom labels for the studies

-   Summary effect sizes for each moderator

## Explanation of the Code

-   The `pdf()` function starts the graphics device driver to create PDF files, and the `file` function specifies the name of the file.

-   The `width` and `height` function adjust the dimensions of the PDF file.

-   The `forest()` function is used to create the forest plot, and the `mlmmetaresults` object contains the results of the meta-analysis.

-   The `rows` function specifies the arrangement of studies by publication type, and in ascending order of effect sizes per publication type.

-   The `ylim` function sets the y-axis limits for the plot.

-   The `ilab` function is used to add sample size information for both presence of smartphones and absence of smartphones groups into the forest plot.

-   The `cbind()` function combines the columns indicating the sample size of the groups (dys and control).

-   The `ilab.xpos` function specifies the horizontal arrangement of the columns, while the `xlim` function sets the x-axis limits for the plot.

-   The `slab` function is used to label each effect size with its respective study, using the `paste()` function to combine the "Paper" and "Study" columns.

-   The `paste()` function creates the label, and the `sep` function specifies the separator between the columns.

-   The `alim` function sets the confidence interval limits, and the `steps` function determines the number of intervals in the x-axis.

-   The `header` function is set to FALSE to allow for manual specification of headers.

-   The `xlab` function specifies the label for the confidence interval, in this case, "Hedge's g".

-   The `text()` function is used to manually include text within the plot, such as the "Author(s) Year" header and specific sample size column headers.

-   The `x` and `y` functions in the `text()` function adjust the position of the headers.

-   The `x` function specifies the horizontal arrangement of the columns and the `y` function specifies the vertical arrangement of the columns.

-   The `font` function adjusts the font size.

-   The `pos` function specifies the position of the text relative to the specified coordinates

-   The `rma.mv()` function is used to perform moderation analysis for each publication type, with the `subset` function specifying the subset of data for each category.

-   The `rest.j`, `res.t`, and `res.c` variables store the results of the moderation analysis for journal articles, thesis/dissertations, and conference papers, respectively.

-   The `subset` function is used to specify the subset of data for categorical moderators, allowing for separate analyses for each category.

-   The `addpoly()` function is used to add summary effect sizes for each of the moderators, with the `row` function specifying the position of the summary in the plot.

-   The `dev.off()` function is used to close the graphics device and finalize the plot as a saved file.

```{r}
### Forest Plot of Moderators --------------

# Save the forest plot as a PDF file
# Name the pdf file of the forest plot
# Adjust the width and height of the pdf file
pdf(file = "mlmforestplotwithmoderators.pdf", width = 15, height = 45) 
forest(
       mlmmetaresults,
       # Manually arrange effect sizes by creativity measure type
       # - Journal article: Rows 143 to 79
       # - Thesis/Dissertations: Rows 75 to 7
       # - Conference: Rows 3 to 2
       # The arrangement must consider spacing and must end at row 2
       rows = c(143:79, 75:7, 3:2),

       # Add y-axis limits
       ylim = c(-3, 147),

       # Add sample size information for presence of smartphones and absence of smartphones groups
       # -4.2 for presence of smartphones (n_p)
       # -3.6 for absence of smartphones (n_a)
       ilab = cbind(n_p, n_a),
       ilab.xpos = c(-4.2, -3.6),

       # Label studies on the forest plot
       slab = paste(author, year_published, sep = ", "),

       # Add x-axis limits
       xlim = c(-7, 4),

       # Add confidence interval limits
       # Adjust intervals based on the number of steps
       alim = c(-1.5, 1.5),
       steps = 11,

       # Change size of effect size polygons
       efac = 0.3,

       # Remove headers (if any), for manual input
       header = FALSE,

       # Add label for confidence interval, in this case, "Hedge's g"
       xlab = "Hedge's g"
)

# For the following lines of code,
# Use text function to manually include text within the plot

# Add text labels for moderator (type of publication)
# Labels for different creativity task types (Moderator Analysis)
# - "Journal article" at y = 7
# - "Thesis/Dissertations" at y = 15
# - "Non-Conference" at y = 21
text(
       x = -7,
       y = c(4, 76, 144),
       pos = 4,
       c("Conference", "Thesis/dissertation", "Journal article"),
       font = 2
)

# Moderation analysis
res.j = rma.mv(
       yi,
       vi,
       random = ~ 1 | lab_id / ID,
       subset = (publication == "Journal article"),
       data = multilevelmeta,
       # To address convergence issues (if it exists)
       control=list(rel.tol=1e-8) 
)
res.t = rma.mv(
       yi,
       vi,
       random = ~ 1 | lab_id / ID,
       subset = (publication == "Thesis/dissertation"),
       data = multilevelmeta
)
res.c = rma.mv(
       yi, 
       vi, 
       random = ~ 1 | lab_id / ID,
       subset = (publication == "Conference"), 
       data = multilevelmeta
)

# Add summary effect sizes for each of the moderators
addpoly(res.c, row = 1) # summary effect for "Conference" group
addpoly(res.t, row = 6) # summary effect for "Thesis/Dissertation" group
addpoly(res.j, row = 78) # summary effect for "Journal article" group

# Add"Author(s) Year" header
text(x = -6.5, y = 146, "Author(s) Year", font = 2) 

# Add “Sample Size” header
text(x = -3.9, y = 146.7, "Sample Size", font = 2)

# Add specific sample size column headers, “Presence” and “Absence”
# x = -4.2 for for presence of smartphones
# x = -3.6 for absence of smartphones
# y = 146 for both
text(c(x = -4.2, x = -3.6), y = 146, c("Presence", "Absence"), font = 2)

# Add "g [95% CI]" header
text(x = 3.6, y = 146, "g [95% CI]", font = 2) 

# Close the forest plot and finalise it as a saved file
dev.off()
```

# End of Code
